---
title: "Revision: Results - FHIR Review"
author: "Lehne, Vorisek"
date: "20 4 2022"
output:
  html_document: default
  word_document: default
  pdf_document: default
editor_options:
  chunk_output_type: console
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



# Network Analysis

First, load relevant packages for the network analysis:
```{r message = FALSE}
library(tidyverse)
library(igraph)
library(ggraph)
library(visNetwork)
library(readxl)
```

## Read in the articles ("articles.xlsx" was exported from Rayyan):

```{r message = FALSE}
library(readxl)
articles <- read_excel("articles.xlsx")
head(articles)
```


Select relevant variables (title and authors of papers), reformat author names, make sure only included articles are considered, correct some author names for consistency, and create variable with list of authors:
```{r}
articles_clean <- articles %>% 
        select(title, authors) %>% 
        mutate(authors = str_remove_all(authors, "[.,]")) %>% 
        mutate(authors = ifelse(authors == "F Khalique and S A Khan",
                                 "Khalique F and Khan SA", authors)) %>% 
        mutate(authors = str_replace_all(authors, " ", "_")) %>% 
        mutate(authors = str_replace_all(authors, "_and_", " ")) %>% 
        mutate(authors = str_replace(authors, "Agrawal_Asha", "Agrawal_A")) %>% 
        mutate(authors = str_replace(authors, "Broadbent_Matthew", "Broadbent_M")) %>%
        mutate(authors = str_replace(authors, "Dobson_Richard_J_B", "Dobson_RJB")) %>%
        mutate(authors = str_replace(authors, "Folarin_Amos", "Folarin_A")) %>%
        mutate(authors = str_replace(authors, "Gale_Darren", "Gale_D")) %>%
        mutate(authors = str_replace(authors, "Gorrell_Genevieve", "Gorrell_G")) %>%
        mutate(authors = str_replace(authors, "Hansen_D", "Hansen_DP")) %>%
        mutate(authors = str_replace(authors, "Hansen_DPP", "Hansen_DP")) %>%
        mutate(authors = str_replace(authors, "Ibrahim_Zina_M", "Ibrahim_ZM")) %>%
        mutate(authors = str_replace(authors, "Jackson_Richard", "Jackson_R")) %>%
        mutate(authors = str_replace(authors, "Kartoglu_Ismail", "Kartoglu_I")) %>%
        mutate(authors = str_replace(authors, "Morley_Katherine_I", "Morley_KI")) %>%
        mutate(authors = str_replace(authors, "Roberts_Angus", "Roberts_A")) %>%
        mutate(authors = str_replace(authors, "Stewart_Robert", "Stewart_R")) %>%
        mutate(authors = str_replace(authors, "Stringer_Clive", "Stringer_C")) %>%
        mutate(authors = str_replace(authors, "Toti_Giulia", "Toti_G")) %>%
        mutate(authors = str_replace(authors, "Wu_Honghan", "Wu_HH")) %>%
        mutate(authors_list = str_split(authors, " "))

head(articles_clean)
```



## Network of authors

Create dataframe with authors and their numbers of publications:
```{r}
# all authors
authors <- unlist(articles_clean$authors_list) %>% 
        unique() %>% 
        data.frame(author = .) %>% 
        mutate(author = as.character(author), publications = 0) %>% 
        arrange(author)

# add number of publications per author
for (i in 1:nrow(authors)) {
        a <- authors$author[i]
        for (j in 1:nrow(articles_clean)) {
                authorlist <- unlist(articles_clean$authors_list[j])
                if (a %in% authorlist) {
                        authors$publications[i] <- authors$publications[i] + 1
                }
        }
}
```

## Get combinations of authors and count their number of co-authorships:
```{r}
# construct dataframe with all combinations of authors
author1 = character()
author2 = character()

for (i in 1:(nrow(authors)-1)) {
        for (j in ((i+1):nrow(authors))) {
                author1 <- c(author1, authors$author[i])
                author2 <- c(author2, authors$author[j])
        }
        
}

# tibble with authors
author_coll <- tibble(author1, author2, collaborations = 0)

# count collaborations for all author combinations
for (i in 1:nrow(author_coll)) {
        
        a1 <- author_coll$author1[i]
        a2 <- author_coll$author2[i]
        
        for (j in 1:nrow(articles_clean)) {
                
                authorlist <- unlist(articles_clean$authors_list[j])
                
                if (a1 %in% authorlist & a2 %in% authorlist) {
                        author_coll$collaborations[i] <- author_coll$collaborations[i] + 1
                }
        }
}
```

## Prepare network object:
```{r}
# edgelist (authors with collaborations)
edgelist_authors <- author_coll %>% 
        filter(collaborations > 0) %>% 
        arrange(desc(collaborations))

# create network object
network_authors <- graph_from_data_frame(edgelist_authors, directed = FALSE,
                                         vertices = authors)
```

### Plot the network:
```{r}
plot_network <- ggraph(network_authors, layout = "kk") +
        geom_edge_link(aes(width = collaborations), color = "grey80") +
        geom_node_point(aes(fill = as.factor(publications), size = publications),
                        shape = 21) +
        #geom_node_point(aes(size = publications)) +
        geom_node_text(aes(label = name), repel = TRUE, size = 2.5,
                       family = "serif") +
        scale_edge_width(range = c(0.5, 2))+
        scale_fill_viridis_d(option = "C") +
        guides(fill = "none", size = "none", width = "none") +
        theme_void() +
        theme(legend.position = "none")

plot_network
```


Each point represents an author. Point size and color indicate the number of publications of this author (between 1 and 6). Lines indicate that authors have co-authored at least one paper together. Line thickness represents the number of co-authorships.

## author information
```{r}
nrow(authors)
summary(authors$publications)
```
```{r}
author_info <- authors %>% 
  group_by(publications) %>% 
  summarize(N = n()) %>% 
  mutate(Percent = N / sum(N) * 100,
         Percent = round(Percent, 1)) %>% 
  arrange(desc(N))

knitr::kable(author_info)
```

***

# Analysis of Excel file

## Data import and cleaning

Import Excel file with included studies ("IncludedStudies2.xlsx"):

```{r message = FALSE}
excel_data <- read_excel("IncludedStudies2.xlsx")
```

Clean data (keep only variables relevant for analysis, rename variables, reformat, ...):

```{r}
excel_data_clean <- excel_data %>% 
  select(Author = `Author (First)`,
         Title,
         Year,
         Country = `Country (according to 1st author affiliation)`,
         Item_mapped_keyword = `Item mapped_keyword`,
         Goal_keyword,
         Other_standards_keyword = `Other standards_keyword`,
         Other_software_keyword = `Other Software_keyword`,
         FHIR_resource_used = `FHIR Resource used`,
         FHIR_extension_used = `FHIR extension used`,
         FHIR_version = `FHIR version`,
         Patients,
         Variables,
         Variables_count,
         Research_category = `Research Category`,
         Medical_area = `Medical Area`,
         Journal,
         Impact_factor = `Impact Factor Research)`,
         Research_area = `Research Area`,
         Clinical_trials = `Clinical Trials`,
         generic,
         SNOMED_CT = `Other standards - SNOMED CT`,
         LOINC = `Other standards - LOINC`,
         ICD_10 = `Other standards - ICD 10`,
         OMOP = `Other standards - OMOP`,
         Other = `Other standards - Other`,
         None = `Other standards - None (nicht angegeben oder nicht verwendet)`) %>% 
  mutate(Journal = str_trim(Journal)) %>% 
  mutate(Journal = ifelse(Journal %in%  c("Studies in Health technology and informatics", "Studies in Health Technology and Informatics"),
                          "Studies in health technology and informatics", Journal)) %>% 
  mutate(Journal = ifelse(Journal == "JMIR medical informatics",
                          "JMIR Medical Informatics", Journal)) %>%
  mutate(Impact_factor = ifelse(Impact_factor %in% c("None", "-"), NA, Impact_factor)) %>% 
  mutate(Impact_factor = str_replace(Impact_factor, ",", ".")) %>% 
  mutate(Impact_factor = round(as.numeric(Impact_factor), 2)) %>% 
  mutate(Country = ifelse(Country == "USA / Switzerland", "USA", Country)) %>% 
  mutate(Research_area = ifelse(Research_area == "Clinical trials",
                               "Clinical Trials", Research_area))
```

## Analysis

### Publications per year
```{r}
# all articles
pub_all <- read_excel("articles_figure1.xlsx")

# number of publications per year (all articles)
pub_all_year <- pub_all %>% 
  group_by(year) %>% 
  summarize(n = n()) %>% 
  filter(year < 2022) %>% 
  mutate(Articles = "all")

# number of publications per year (included articles)
pub_included_year <- excel_data_clean %>% 
  group_by(Year) %>% 
  summarize(n = n()) %>% 
  filter(Year < 2022) %>% 
  mutate(Articles = "included") %>% 
  rename(year = Year)

# combine in one data frame
pub_year <- rbind(pub_all_year, pub_included_year)
pub_year$year <- as.integer(pub_year$year)

# plot
plot_pub <- ggplot(pub_year) + geom_line(aes(year, n, linetype = Articles),
                             color = "#1b3051") +
  labs(x = "", y = "Number of Publications", linetype = "") +
  scale_y_log10(minor_breaks = c(seq(1,9), seq(10,150,10))) +
  theme(rect = element_blank(),
        axis.line = element_line(color = "grey40"),
        panel.grid.major = element_line(color = "grey90"),
        panel.grid.minor.y = element_line(color = "grey90"))

plot_pub + scale_x_continuous(limits = c(2013, 2021), breaks = c(2013, 2015, 2017, 2019, 2021))

```

Summary publications per year
```{r}
summary(excel_data_clean$Year)
```



### Countries

The articles originate from the following countries (according to affiliation of first author):

```{r}
countries <- excel_data_clean %>% 
  group_by(Country) %>% 
  summarize(N = n()) %>% 
  mutate(Percent = N / sum(N) * 100,
         Percent = round(Percent, 1)) %>% 
  arrange(desc(N))

knitr::kable(countries)
```

```{r}
# define theme and colors for barplots
bar_theme <-   theme(rect = element_blank(),
        axis.ticks.y = element_line(color = "grey40"),
        panel.grid.major.x = element_line(color = "grey90"))
bar_color <-  "#1b3051"
bar_width <- 0.6

plot_country <- ggplot(countries) +
  geom_col(aes(reorder(Country, N, sum), N), fill = bar_color,
           width = bar_width) +
  labs(x = "", y = "Number of studies") +
  scale_y_continuous(breaks = seq(0,30,5)) +
  coord_flip() +
  scale_color_manual(values = "#668dcc") +
  bar_theme

plot_country

```


### Journals

The articles were published in the following scientific journals:

```{r}
journals <- excel_data_clean %>% 
  group_by(Journal) %>% 
  mutate(N = n()) %>% 
  select(Journal, Impact_factor, N) %>% 
  distinct() %>% 
  ungroup() %>% 
  mutate(Percent = N / sum(N) * 100,
         Percent = round(Percent, 1)) %>% 
  arrange(desc(N))

knitr::kable(journals)
```


### Research category

Articles were grouped into the following research categories:

```{r}
categories <- excel_data_clean %>% 
  group_by(Research_category) %>% 
  summarize(N = n()) %>% 
  mutate(Percent = N / sum(N) * 100,
         Percent = round(Percent, 1)) %>%
  arrange(desc(N))

knitr::kable(categories)
```

```{r}
plot_category <- ggplot(categories) + geom_col(aes(reorder(Research_category, N, sum), N),
                              fill = bar_color,
                              width = bar_width) +
  labs(x = "", y = "Number of studies") +
  scale_y_continuous(breaks = seq(0,20,5)) +
  coord_flip() +
  bar_theme

plot_category
```

Most articles reported generic solutions:

```{r}
table(excel_data_clean$generic)
```

### Medical area

Articles were grouped into the following medical areas:

```{r}
medical_area <- excel_data_clean %>% 
  group_by(Medical_area) %>% 
  summarize(N = n()) %>% 
  mutate(Percent = N / sum(N) * 100,
         Percent = round(Percent, 1)) %>%
  arrange(desc(N))

knitr::kable(medical_area)
```

```{r}
plot_medical <- ggplot(medical_area) + geom_col(aes(reorder(Medical_area, N, sum), N),
                                fill = bar_color,
                                width = bar_width) +
  labs(x = "", y = "Number of studies") +
  scale_y_continuous(breaks = seq(0,30,5)) +
  coord_flip() +
  bar_theme

plot_medical
```

### Research area

Articles were grouped into the following research areas:

```{r}
research_area <- excel_data_clean %>% 
  group_by(Research_area) %>% 
  summarize(N = n()) %>% 
  mutate(Percent = N / sum(N) * 100,
         Percent = round(Percent, 1)) %>%
  arrange(desc(N))

knitr::kable(research_area)
```

```{r}
plot_area <- ggplot(research_area) +
  geom_col(aes(reorder(Research_area, N, sum), N),
           fill = bar_color,
           width = bar_width) +
  labs(x = "", y = "Number of studies") +
  scale_y_continuous(breaks = seq(0,30,5)) +
  coord_flip() +
  bar_theme

plot_area
```

## Plot clinical trials as part of clinical research:
```{r}
plot_data <- excel_data_clean %>% 
  mutate(Clinical_trials = ifelse(Research_area == "Clinical Trials",
                                  "Clinical Trials", "Other")) %>% 
  mutate(Research_area = ifelse(Research_area == "Clinical Trials",
                                "Clinical Research", Research_area)) %>% 
  mutate(Research_area = ifelse(Research_area == "Public Health/ Epidemiological Studies", "Public Health/ \n Epidemiological\n Studies", Research_area))

plot_clinical <- ggplot(plot_data) +
  geom_bar(aes(Research_area,
               fill = Clinical_trials), width = bar_width) +
  labs(x = "", y = "Number of studies", fill = "") +
  scale_y_continuous(breaks = seq(0,35,5)) +
  scale_fill_manual(values=c("#668dcc", "#1b3051")) +
  coord_flip() +
  bar_theme

plot_clinical
```

### Standards

The following other standards and terminologies were used in the studies:

```{r}
standards <- excel_data_clean %>% 
  select(SNOMED_CT:None) %>% 
  gather("Standard", "used") %>% 
  group_by(Standard) %>% 
  summarize(N = sum(used)) %>% 
  mutate(Percent = N / nrow(excel_data_clean) * 100,
         Percent = round(Percent, 1)) %>%
  arrange(desc(N))

knitr::kable(standards)
```

```{r}
plot_standards <- ggplot(standards) + geom_col(aes(reorder(Standard, N, sum), N),
                             fill = bar_color,
                             width = bar_width) +
  labs(x = "", y = "Number of studies") +
  scale_y_continuous(breaks = seq(0,21,5)) +
  coord_flip() +
  bar_theme

plot_standards
```

### Goals

This section analyzes the most common keywords describing the goals of the articles.


Load tidytext package and get word frequencies of all words in Goal_keyword variable:
```{r}
library(tidytext)

# define some stop words to be excluded
stop_words <- c("of", "and", "for", "from", "a", "an", "between",
                "across", "on", "the", "to", "in", "into", "with",
                "as", "or", "other", "such", "via")

goal_keywords <- excel_data_clean %>% 
  select(Goal_keyword) %>% 
  unnest_tokens(word, Goal_keyword, to_lower = TRUE) %>% 
  filter(!(word %in% stop_words)) %>% 
  group_by(word) %>% 
  summarize(freq = n()) %>% 
  arrange(desc(freq))

```

The wordcloud shows the most common words mentioned as keywords:
```{r}
library(wordcloud)
```

```{r}
wordcloud(words = goal_keywords$word, freq = goal_keywords$freq, min.freq = 1,                      max.words=150, random.order=FALSE, rot.per=0.10, 
             colors=c("#335a99", "#1b3051"))

```


# Create Images for Publication out of multiple graphs
Package gridExtra
```{r}
library(gridExtra)
```

```{r}
require(gridExtra)
plot1 <- ggplot(categories) + geom_col(aes(reorder(Research_category, N, sum), N),
                              fill = bar_color,
                              width = bar_width) +
  labs(x = "", y = "Number of studies") +
  scale_y_continuous(breaks = seq(0,20,5)) +
  coord_flip() +
  bar_theme
plot2 <- ggplot(standards) + geom_col(aes(reorder(Standard, N, sum), N),
                             fill = bar_color,
                             width = bar_width) +
  labs(x = "", y = "Number of studies") +
  scale_y_continuous(breaks = seq(0,21,5)) +
  coord_flip() +
  bar_theme
```


```{r}
test <- grid.arrange(plot1, plot2, ncol=2)

test

ggsave("test.png", arrangeGrob(plot1, plot2))

```

package cowplot
```{r}
library(cowplot)

```
```{r}
p2 <- plot_grid(plot_medical, plot_category, plot_area, plot_standards, labels = "AUTO")
p2
```

## Save Images for Publication
```{r}
ggsave("Figure5test.png", p2)
ggsave("Figure2.png", plot_pub)
ggsave("Figure3.png", plot_network)
ggsave("Figure4A.png", plot_country)
ggsave("Figure5A.png", plot_medical)
ggsave("Figure5B.png", plot_category)
ggsave("Figure5C.png", plot_clinical)
ggsave("Figure5D.png", plot_standards)
```
